Bayesian multi-proxy reconstruction of latitudinal temperature gradients for the Early Eocene

### 2023-01-23 manuscript/manuscript.Rmd
results

# sdy
eeco_sdy <- readRDS("results/eeco/eeco_sdy.rds")
modern_eeco_sample_sdy <- readRDS("results/modern/modern_eeco_sample_sdy.rds")

@@

## EECO reconstruction
<!-- - Show Eocene gradient
- Global t : use raster package for calculation of areal weights
- report lat gradient
-->

The modelled Eocene temperature gradient is starkly different from the modern (Fig 4). Modelled, median equatorial temperatures are `r tdiff_lat0$"50%"` (95% CI: `r tdiff_lat0$"2.5%"` - `r tdiff_lat0$"97.5%"`) degC higher for the Eocene, and polar temperatures are higher by `r tdiff_lat90$"50%"` (95% CI: `r tdiff_lat90$"2.5%"` - `r tdiff_lat90$"97.5%"`) degC. This results in a strongly flattened latitudinal temperature gradient of `r eocene_gradient$"50%"`  (95% CI: `r eocene_gradient$"2.5%"` - `r eocene_gradient$"97.5%"`) degC for the Eocene, as opposed to `r modern_gradient$"50%"` degC for the modern.

The high variability of early Eocene palaeotemperature proxies, particularly in the mid-latitudes, and the scarcity of high-latitude data, results in substantial uncertainties in the modeled temperature gradient. This is reflected in the residual standard deviation ($\sigma$) of the early Eocene gradient, `r eeco_sdy$"50%"` (95% CI: `r eeco_sdy$"2.5%"` - `r eeco_sdy$"97.5%"`) degC, more than twice as high than  $\sigma$ of the modern gradient, resampled at early Eocene latitudes, `r modern_eeco_sample_sdy$"50%"` (95% CI: `r modern_eeco_sample_sdy$"2.5%"` - `r modern_eeco_sample_sdy$"97.5%"`) degC. The inclusion of ecological proxy data only slightly reduces the overall model standard deviation, but helps constrain polar temperature estimates (Supplementary materials).  ### Continue here.

@@

>>>>

# sdy
eeco_sdy <- readRDS("results/eeco/eeco_sdy.rds")
modern_eeco_sample_sdy <- readRDS("results/modern/modern_eeco_sample_sdy.rds")
modern_sdy <- readRDS("results/modern/modern_sdy.rds")

@@

## EECO reconstruction

The modeled Eocene temperature gradient is starkly different from the modern (Fig 4). Modelled, median equatorial temperatures are `r tdiff_lat0$"50%"` (95% CI: `r tdiff_lat0$"2.5%"` - `r tdiff_lat0$"97.5%"`) degC higher for the Eocene, and polar temperatures are higher by `r tdiff_lat90$"50%"` (95% CI: `r tdiff_lat90$"2.5%"` - `r tdiff_lat90$"97.5%"`) degC. This results in a strongly flattened latitudinal temperature gradient of `r eocene_gradient$"50%"`  (95% CI: `r eocene_gradient$"2.5%"` - `r eocene_gradient$"97.5%"`) degC for the Eocene, as opposed to `r modern_gradient$"50%"` degC for the modern.

The high variability of early Eocene palaeotemperature proxies, particularly in the mid-latitudes, and the scarcity of high-latitude data, result in substantial uncertainties in the modeled temperature gradient. This is reflected in the residual standard deviation ($\sigma$) of the early Eocene gradient, `r eeco_sdy$"50%"` (95% CI: `r eeco_sdy$"2.5%"` - `r eeco_sdy$"97.5%"`) degC, more than twice as high than  $\sigma$ of the modern gradient, `r modern_sdy$"50%"`.

@@

For comparison, a model run excluding the ecological proxies results in a global mean temperature of  
The inclusion of ecological proxy data only slightly reduces the overall model standard deviation (Supplementary materials).
